---
title: "Untitled"
output: html_document
---

Learning to use the Recipes tidymodels package
- helps preprocess data before training your model such as 
    - convert qualitative predictors to indicator variables
    - transform data to be on different scale 
    - transform groups of predictors together
    - extract key features from raw variables (like day of week from date variable)

```{r}
library(tidymodels) #for recipe package

#helper packages
install.packages("nycflights13")
library(nycflights13) #for flight data

install.packages("skimr")
library(skimr) #for variable summaries
```

```{r}
summary(flights)
```


```{r}
set.seed(123)


flight_data <- 
  flights %>%
  mutate(
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    #labeling those 30 or more minutes late as late and all others on time
    arr_delay = factor(arr_delay),
    #convert the arrival delay to a factor with two levels (late, or on time)
    date = lubridate::as_date(time_hour)
    #code that makes looking at dates easier, takes the time/hour/date data from the original data set and changes it to only the date
  )%>%
    #include weather data 
  inner_join(weather, by = c("origin", "time_hour")) %>%
  #only retain the specific columns we will use (excludes several columns)
  select(dep_time, flight, origin, dest, air_time, distance, carrier, date, arr_delay, time_hour) %>%
  #exclude missing data
  na.omit() %>% 
  #better to have qualitative columns encoded as factors instead of character strings - factors use values and character strings use letters
  mutate_if(is.character, as.factor)
```

```{r}
summary(flight_data)
```


```{r}
#determine the percent of flights that arrived 30 or more minutes late

flight_data %>%
  count(arr_delay)%>%
  mutate(prop = n/sum(n))

```

```{r}
glimpse(flight_data) #beware that we do not want to use flight or time_hour as a predictor in our model 
```


```{r}
flight_data %>% 
  skimr::skim(dest, carrier) #to determine the number of flight destinations and the number of distinct carriers (brands like Delta etc) - skim is an alternate to summary 

#will be using simple logistic regression model so destination and carrier will be converted to dummy variables - set of artificial numbers that represent categorical data (like using binary code to represent the days of the week)
```


Split the data in 2 
```{r}
#splitting dataset into 2 parts - training data to FIT the model, testing data to measure model performance

set.seed(222) # fix the random numbers which allows this to be reproducible

#putting 3/4 of the data into the training set 
data_split <- initial_split(flight_data, pop = 3/4)

#create data frames from both 
train_data <- training(data_split)
test_data <- testing(data_split) 
```


Create recipe and roles
```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data)%>%
#recipe function has 2 arguments: formula (any variable on the left side of the ~ would be considered the model outcome (aka arr_delay). on the right side would be the predictors or you can use the dot to indicate all other variables as predictors) - OR data (recipe is associated with the data set used to create the model, usually used for the training set which is why data is train_data here)

update_role(flight, time_hour, new_role = "ID") #let recipes know that flight and time_hour are variables with a custom role called "ID" - tells the recipe to keep these two variables but not use them as outcomes or predictors in the model - "ID" can be replaced with "predictor" or "outcome" to change the role in the model

summary(flights_rec)

```

Create features
```{r}
#maybe date has something to do with the likelihood of being late, we need to change date into a numeric value

flight_data %>% 
  distinct(date) %>% 
  mutate(numeric_date = as.numeric(date))

```


```{r}
#however, your model may be more complete by being able to see the day of the week, the month, and whether or not the date corresponds to a holiday 

flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) %>%
  update_role(flight, time_hour, new_role = "ID")%>%
  step_date(date, features = c("dow", "month")) %>% #adds two new factor columns, day of the week and the month 
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"),
               keep_original_cols = FALSE) #binary variable listing holiday or not - the keep_original_cols means we are removing the original                date variable since we no longer need it in the  model

```


Since we are creating a logistic regression model, the predictors need to be numeric instead of nominal (string and factor)

#you have to tell the model to change predictors into numerical value in this case because not all models require numeric predictors 

```{r}
#making the categorical (aka nominal) data into dummy variables (assign binary values to each)
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) %>%
  update_role(flight, time_hour, new_role = "ID")%>%
  step_date(date, features = c("dow", "month")) %>% 
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"),
               keep_original_cols = FALSE)%>%
  step_dummy(all_nominal_predictors()) #applied to all nominal predictors instead of isolation an individual variable (selects the origin, dest, and carrier, date_dow, and date_month)


```

```{r}
#check to see if there are any destinations or origins present in the test set and not the training set

test_data %>%
  distinct(dest) %>%
  anti_join(train_data)

test_data %>%
  distinct(origin) %>%
  anti_join(train_data)

#find that there is one destination that is present in the test data but not the training data, and no origins
#we will use step_zv() in order to remove columns from the data when the training set have a single value. needs to be added AFTER step_dummy
```


```{r}
flights_rec <- 
  recipe(arr_delay ~ ., data = train_data) %>%
  update_role(flight, time_hour, new_role = "ID")%>%
  step_date(date, features = c("dow", "month")) %>% 
  step_holiday(date, 
               holidays = timeDate::listHolidays("US"),
               keep_original_cols = FALSE)%>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

#now the recipe is complete! how do we use it? 
```


Merging the recipe and the model 

```{r}
#building a model specification with the parsnip package

lr_mod<- 
  logistic_reg()%>%
  set_engine("glm")

#bundle the model and recipe together using the workflows package (makes it easier to train and test workflows)

flights_wflow <-
  workflow()%>%
  add_model(lr_mod) %>%
  add_recipe(flights_rec)


#prepare the recipe and train the model from the resulting predictors 
flights_fit <-
  flights_wflow %>%
  fit(data = train_data)

#recipe is now finalized and model objects were fitted inside 

```

```{r}

#if you want to extract the model or recipe objects from the workflow: 
flights_fit %>% 
  extract_fit_parsnip()%>%
  tidy()

flights_fit %>%
  extract_recipe() %>%
  tidy()
```


Now we need to use the workflow (flights_fit) with the test data 
```{r}
predict(flights_fit, test_data)

#this returns the prediction for if each flight will be on time or late (by 30 mins or more)
```


If we want to know the probability of a flight being on time instead, we can do this: 
```{r}
#save the model and test data together
flights_aug <- 
  augment(flights_fit, test_data)

flights_aug %>% 
  select(arr_delay, time_hour, flight, .pred_class, .pred_on_time)
```


Now we want to know the accuracy of our model (calculate a metric that tells us how well the model predicted late arrivals as compared to the actual data in column "arr_delay")

```{r}
#The area under an ROC (receiver operating characteristic) curve can help us determine this - frequently used method to evaluate the performance of binary algorithms

flights_aug %>%
  roc_curve(truth = arr_delay, .pred_late) %>%
  autoplot() #makes the ROC plot

flights_aug %>%
  roc_auc(truth = arr_delay, .pred_late) #calculates area under the curve
```













Trying the model without the recipe now just to see if the recipe improved it at all - I don't think I did this right haha. Someone let me know if you see what I did wrong. 

```{r}
flight_data_2 <- 
  flights %>%
  mutate(
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    arr_delay = factor(arr_delay),
    date = lubridate::as_date(time_hour)
  )%>%
  inner_join(weather, by =  c("origin", "time_hour")) %>%
  select(dep_time, origin, dest, air_time, distance, carrier, date, arr_delay) %>%
  na.omit() %>% 
  mutate_if(is.character, as.factor)

flights_data_2
```


```{r}
set.seed(84) 
data_split_2 <- initial_split(flight_data_2, pop = 3/4)


train_data_2 <- training(data_split_2)
test_data_2 <- testing(data_split_2) 

test_data_2 %>%
  distinct(dest) %>%
  anti_join(train_data_2)

```

```{r}
lr_mod_2<- 
  logistic_reg()%>%
  set_engine("glm")

lr_fit_2 <-
  lr_mod_2 %>%
  fit(arr_delay ~ date * dest, data = train_data_2)

```


```{r}
flights_aug_2 <- 
  augment(lr_fit_2, test_data_2)

flights_aug_2 %>% 
  select(arr_delay,.pred_class, .pred_on_time)
```


```{r}
flights_aug_2 %>%
  roc_curve(truth = arr_delay, .pred_late) %>%
  autoplot() #makes the ROC plot

flights_aug_2 %>%
  roc_auc(truth = arr_delay, .pred_late) #calculates area under the curve
```

Shows that the results are much less accurate without using the recipe (if I did this correctly).





