---
title: "Tidyverse Tutorial 3"
output: html_document
---

Focuses on how to characterize model performance based on resampling statistics (drawing repeated samples from the original data - creates repetitive cases to determine accuracy)

```{r}
install.packages("ranger")
install.packages("modeldata")
install.packages("tidymodels")

library(tidymodels)

library(modeldata) #provides the cell data needed for the tutorial (from Hill et al. 2007)
```

```{r}
data(cells, package = "modeldata")

cells #58 variables, the one most valuable to us is "class" 

```


```{r}
cells %>% 
  count(class)%>%
  mutate(prop = n/sum(n))

```

```{r}
set.seed(123) #(keeps it reproducible)

cell_split <- initial_split(cells %>% select(-case),strata = class) #select(-case) means that the case column is left out when creating our split, and using the strata argument makes a stratified split which means that the training and testing set will contain approximately even proportions of PS and WS (helpful when we have such uneven data, as seen above) 

```

```{r}
cell_train <- training(cell_split)
cell_test <- testing(cell_split)

nrow(cell_train)

nrow(cell_train)/nrow(cells) 
```

```{r}
#Now we want to compare the training set class proportions vs test set class proportions to make sure they are even (remember that in the cell dataset there was approximately 64% PS and 36% WS)

cell_train %>% 
  count(class) %>%
  mutate(prop = n/sum(n))


cell_test %>% 
  count(class) %>% 
  mutate(prop = n/sum(n))

#we can see that in both the testing and training data, the PS class makes up approximately 64%. Great!
```

















Creating the model 
```{r}
rf_mod <-
  rand_forest(trees = 1000)%>% 
  set_engine("ranger")%>% 
  set_mode("classification") 
```


Fitting the model to the data 
```{r}
set.seed(234) 

rf_fit <-
  rf_mod %>%
  fit(class~ ., data = cell_train) 

rf_fit
```


Estimating performance with train data 
```{r}
rf_training_pred <- predict(rf_fit, cell_train) %>% 
  bind_cols(predict(rf_fit, cell_train, type = "prob")) %>%
  bind_cols(cell_train %>% select(class)) 

rf_training_pred

rf_training_pred %>% 
  roc_auc(truth = class,.pred_PS) 

rf_training_pred%>% 
  accuracy(truth=class,.pred_class) 
```

Estimating performance with test data now 
```{r}
rf_testing_pred <- predict(rf_fit, cell_test) %>% 
  bind_cols(predict(rf_fit, cell_test, type = "prob")) %>%
  bind_cols(cell_test %>% select(class))


rf_testing_pred %>%
  roc_auc(truth = class, .pred_PS)

rf_testing_pred %>% 
  accuracy(truth = class, .pred_class) 
```





















Resampling!

Fit a model with resampling
```{r}
set.seed(345)

folds <- vfold_cv(cell_train, v=10) 

folds
```

Building an object for resampling
```{r}
rf_wf <-
  workflow() %>%
  add_model(rf_mod) %>%
  add_formula(class ~ .)

set.seed(456)
rf_fit_rs <- 
  rf_wf %>%
  fit_resamples(folds) 

rf_fit_rs
```

```{r}
collect_metrics(rf_fit_rs)
```



